### 2019-02-15 tex_v2/manuscript.tex
got the kraken, updating everyhting

\usepackage{amsmath}
\usepackage{graphicx}
%TODO remove package: blindtext
\usepackage{blindtext}
\usepackage{gensymb}
\usepackage{textgreek}
\usepackage{xcolor}
\usepackage[colorlinks]{hyperref}
	\hypersetup{colorlinks,
	linkcolor={red!50!black},
	citecolor={blue!60!black},
	urlcolor={blue!40!black}
	}
\usepackage{cleveref}
\graphicspath{{figures/}}

%TODO NOTES! REMOVE FROM COMPLETE DRAFT-----------------------------------------
\usepackage{xargs}
\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
\newcommandx{\unsure}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
\newcommandx{\change}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
\newcommandx{\info}[2][1=]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,#1]{#2}}
\newcommandx{\improvement}[2][1=]{\todo[linecolor=purple,backgroundcolor=purple!25,bordercolor=purple,#1]{#2}}
\newcommandx{\noshow}[2][1=]{\todo[disable,#1]{#2}}
%-------------------------------------------------------------------------------

\begin{document}
	%word count: 156
	\begin{abstract}
		The acorn weevil snout exhibits remarkable flexibility and toughness that are derived from from the microarchitecture of its exoskeleton.
		Here we characterize modifications to the composite profile of the rostral cuticle that simultaneously enhance the flexibility and fracture toughness of the distal portion of the snout.
		Using Classical Laminate Plate Theory (CLPT) we estimate the effect of these modifications on the elastic behavior of the exoskeleton.
		Additionally, changes in the relative layer thicknesses and orientation angles of layers in the exoskeleton are related to the tensile behavior of the snout in six species of diverse morphology, and we demonstrate that a highly curved rostrum can be completely straightened without structural damage.
		We demonstrate that increased endocuticle thickness is strongly correlated with increased tensile strength in the snout.
		Consequently, snout stiffness is shown to be inversely correlated with fracture toughness.
		Finally, we identify exocuticle rich invaginations of the occipital sutures both as a likely site of crack initiation in tensile failure and as a source of morphological constraint on the evolution of the snout.

@@

	%title and author block
	{\title{Exoskeletal microstructure and tensile behavior of the acorn weevil rostrum}
	%CONSIDER CHANGING THE TITLE AFTER WRITING MANUSCRIPT
	\date{\today}

@@

	{The exoskeleton of Coleoptera (beetles) is a hierarchically-structured fibrous composite characterized by variously arranged \textalpha-chitin (N-acetylglucosamine) nanofibrils embedded in a heterogeneous protein matrix \cite{Hepburn1973, Jansen2016, Vincent2004}.
	Although \textalpha-chitin is brittle and strongly anisotropic, beetle cuticle is simultaneously rigid and tough due to its unique laminate microstructure (see kamp, vincent for review), which we characterize in detail below \cite{Kamp2010, Kamp2015, Neville1976}.
	Impact-prone areas and exaggerated structures in arthropods generally exhibit cuticle organization that resists deformation and fracture (e.g. \cite{Amini2015, Mccullough2014mech, Mccullough2014struc, Dirks2012, Dirks2013}); however, acorn weevils in the genus \textit{Curculio} Linnaeus, 1758\footnote{
		Pursuant to the International Code of Zoological Nomenclature, the first mention of any specific epithet will include the full genus and species names as a binomen (two part name) followed by the author and date of publication of the name.
		This is not an in-line reference; it is a part of the name itself and refers to a particular species-concept as indicated in the description of the species by that author \cite{ICZN1999}.}
	instead exhibit unusual distal flexibility in an elongate extension of the head called the rostrum (snout) \cite{Toju2005, Jansen2016, Singh2016, Gibson1969}.
	The rostrum is a hollow, strongly curved (over 90$\degree$ in some species), cylindrical, exoskeletal extension of the otherwise nearly-spherical head, which bears at its apex the terminal chewing mouthparts \cite{Morimoto2003, Ting1933, Ting1936, Dennell1942, Gibson1969}.
	This structure is used by the female to feed and to excavate sites for egg-laying (oviposition, see Fig.~1c), and can be repeatedly bent without evident damage despite being composed of the same material as other rigid body parts \cite{Jansen2016, Singh2016, Gibson1969, Toju2005}.
	By maintaining constant pressure on the snout and rotating around the bore-hole, females are able to flex the rostrum into a straightened configuration and produce a linear channel into the fruit; a single female may prepare hundreds of such sites as an adult \cite{Toju2005, Moffett1989, AguirreUribe1978}.

@@

	We additionally describe the fracture mechanics of the snout and consider how modification of the cuticle may prevent crack formation during oviposition.
	Based on our findings, we posit that the composite profile of the rostral apex enables the snout to undergo repeated distal flexion while remaining within the elastic limits of the material, mitigating the risk of structural damage, and without evident alteration of the mechanical properties of the individual components of the cuticle across the structure and between species.
	Thus, the flexibility and tensile strength of the rostrum appear to be derived \emph{exclusively} from modification of the composite architecture of the exoskeleton.
	To our knowledge, this is the first time that a modified composite profile has been reported as a means of enhancing structural elasticity in the insect exoskeleton (but see \cite{Matsumura2017}).
}	
	\section{Microstructure of the \textit{Curculio} rostrum}
		In arthropods (including beetles), the exocuticle is comprised of numerous unidirectional laminae of chitin nanofibrils; each layer is the thickness of a single fiber (2-4\,nm) embedded in a proteinaceous matrix \cite{Nikolov2010, Nikolov2011}.
		These layers are stacked at a more or less constant angle to each other, forming a quasi-isotropic laminate known as the Bouligand structure \cite{Blackwell1980, Bouligand1972, Neville1976}. 
		This layout effectively mitigates the strong anisotropy of \textalpha-chitin to yield a versatile building material for the exoskeleton \cite{Vincent1982, Vincent2004, Nikolov2010, Nikolov2011}.		
		Beetle endocuticle, however, is unique among arthropods and is comprised of large (1-5\,{\textmu}m diameter in \textit{Curculio}) unidirectional bundles of chitin, called macrofibers.
		Chitin macrofibers are orthotropic (axial: $E_1~=~8.5\,\text{GPa}$, transverse: $ E_2~=~E_3~=~0.52\,\text{GPa}$ \cite{Jansen2016}) and arranged in unidirectional plies, seen in Figs.~REF \cite{Kamp2010, Kamp2015}.
		Typically, adjacent macrofiber laminae are paired and pseudo-orthogonal (i.e., angled approx. $90\degree$ to each other \cite{Cheng2009}, see Figs.~REF), with a constant stacking angle \emph{between} pairs, although other configurations have been observed \cite{Hepburn1973, Kamp2010, Kamp2015, Leopold1992}.
		This geometric sequence of the macrofiber laminae yields an approximately transversely isotropic composite, similar to the Bouligand structure \cite{Kamp2015, Nikolov2010}.
		Notably, the resulting laminate is less rigid than the exocuticle, but exhibits greater toughness because the pseudo-orthogonal plies effectively inhibit crack formation and propagation between successive layers \cite{Kamp2010, Kamp2015, Hepburn1973}.

		Serial thin sectioning and scanning electron microscopy of fractured \textit{Curculio} specimens revealed that endocuticle in the head capsule fits this general profile, with an angle of approximately $30\degree$ between successive pairs of pseudo-orthogonal plies.
		Additionally, in the head capsule, the thickness of the exocuticle and endocuticle in cross section are nearly equal (typically between 20-30\,{\textmu}m).
		However, we found that the cuticle composite lay-up of the rostral apex is strongly differentiated from the head capsule (see Fig.~REF).
		Distally the exocuticle is reduced to a thin shell (ca. 5\,{\textmu}m), with the endocuticle thickened to offset this reduction and maintain a constant cuticle thickness (ca. 50\,{\textmu}m total) throughout its length.
		Moreover, the endocuticular macrofibers exhibit no rotation between successive pseudo-orthogonal plies, and are oriented at approximately $\pm45\degree$ to the longitudinal axis of the snout (i.e., an antisymmetric $[\pm45\degree]$ angle-ply laminate).
		We previously identified these modifications to the composite structure of the cuticle within a single species, \textit{C.~longinasus} Chittenden, 1927 \cite{Jansen2016, Singh2016}; however this composite profile has now been uncovered in the rostrum of six additional, phylogenetically disparate, species (detailed in methods), indicating that this is likely a genus-wide trait.
		In all examined species, the portion of the snout between the head capsule and apex of the scrobe exhibits a gradual transition in composite profile along an anterior-posterior gradient.

@@

	\section{Force-controlled loading to fracture}
		To better characterize the failure behavior of the rostrum, we performed tensile testing on the snouts of six \textit{Curculio} species that representing a mixture of closely and distantly related taxa (see Methods) \cite{Hughes2004phylo, Hughes2004eco, Bonal2016, Bonal2011}.
		Each specimen was immersed in $\text{di-H}_2\text{O}$ for 24 hours, to simulate the living tissue (see \cite{Klocke2011}), then subjected to force-controlled, uniaxial loading to fracture at a constant stress rate of $1.0\,\text{gf}\cdot\text{s}^{-1}$ (detailed in Methods).
		In general, the specimens exhibited a non-linear viscoelastic response curve characterized by a sharp increase in stress at higher strains, terminating in brittle fracture \cite{Mihai2017}.
		We postulate that strain hardening occurs as the longitudinal axis of the macrofibers becomes more closely aligned to the cylindrical axis of the rostrum, thereby resisting tension more directly with increasing strain \cite{Munster2013}.

		We examined the correspondence between composite structure and mechanical behavior of the snout using phylogenetic linear mixed effects models (PGLMM) to account for phylogenetic non-independence in residual variance, with species membership included as a random effect \cite{Revell2010, Felsenstein1985}.
		Our models show that the maximum force sustained at the site of failure was strongly correlated with the cross-sectional area of the endocuticle, and not the exocuticle, at that site.
		Accordingly, there was a negative correlation between ultimate tensile strength of the specimen and the ratio of exocuticle to endocuticle cross-sectional area at the site of fracture.
		Although CLPT predicts a positive association between the proportion of exocuticle and stiffness of a generalized cuticle, we found no correspondence between the cross-sectional properties of the fracture site and the gross behavior of the entire rostrum.
		Because the cross-sectional areas of the cuticle regions vary across the length of the head along an anterior-posterior gradient, it is not possible to correlate measurements from the fracture surface to the properties of the entire snout.

		Instead, we found that the uniaxial elastic modulus (low strain: $E_{low}$) and secant modulus at failure ($E_{sec}$) were inversely correlated with ultimate strain and fracture toughness, implying that stiffer specimens, and by extension, stiffer cuticle profiles, are generally more brittle.
		We also observed a moderate, stiffening size-effect with respect to rostral length, contrary to our expectation that a longer, more strongly curved (see \cite{Hughes2004eco, Bonal2011}) snout would require increased flexibility to avoid fracture during oviposition.
		One possible explanation for this trend is that longer snouts may have a longer transition gradient from basal to apical profile to reinforce the junction between the rostrum and head capsule against buckling.
		Thus, Young's modulus of the rostrum would be comparatively higher in these species because of the higher volume fraction of exocuticle.
		From these tests, we infer that the gross elastic behavior of the cuticle is consistent across the genus, and that a single mechanism (i.e. the modified composite profile) confers increased flexibility and tensile strength to the rostral apex in the genus \textit{Curculio}.
		In addition, the endocuticle demonstrably contributes more to rostral tensile strength than the exocuticle, likely because of its organization into large bundles of aligned, anisotropic fibers, leading to a trade-off between rigidity and toughness.
		Consequently, the altered composite profile of the cuticle in the rostral apex makes the rostrum simultaneously more flexible and fracture resistant.

	\section{Load cycling of \textit{Curculio caryae}}
		To confirm that repeated, proscribed straightening of the rostrum does not result in damage to the cuticle, we performed displacement-controlled fatigue testing on a typical female specimen of \textit{Curculio caryae}, a species that exhibits extreme ($80-90\degree$) rostral curvature \cite{AguirreUribe1978, Gibson1969}.
		The specimen was aligned so that uniaxial tension would induce elongation of the distal portion of the rostrum with minimal off-axis deflection of the un-curved section.
		The strain per cycle was fixed at an amplitude sufficient to completely elongate the snout and generate a tensile load of 1.0\,N  in the straightened configuration (ca. 20\% ultimate strength) at a frequency of $0.33\,\text{Hz}$.
		The test was terminated after a period of two weeks (ca.~400K cycles) when the stress amplitude appeared to reach an asymptotic minimum.		
		The rostrum behaved viscoelastically, indicated by histeresis in the stress-strain relationship during each cycle.
		Strain amplitude decreased logarithmically with cycle number, and the specimen initially appeared to have deformed plastically during the test.
		We believed that this indicated damage to the specimen; however, after cleaning the specimen in a 24 hour wash with ethanol and water, we observed that rostrum returned to its original shape.

@@

	\section{Fractography of test specimens}
		Although it was not always possible to identify void nucleation and crack initiation sites given the complex failure modes evident in the fractured specimens, we observed several patterns characteristic of both the micro-scale behavior of the cuticle and the meso-scale behavior of the rostrum during uniaxial tensile failure, which we describe here (see \cite{Greenhalgh2009} for general approach).

		In transverse view, the exocuticle consistently presented a nearly continuous fracture surface, characteristic of comparatively brittle failure, presumably due to the relatively homogeneous arrangement of \textalpha-chitin laminae in the Bouligand structure \cite{Nikolov2010, Nikolov2011}.
		The exocuticle typically appeared to fracture at lower strains than the endocuticle, with shear-cusp formation evident both at the fracture surface and across exocuticle (but with endocuticle intact) adjacent to the plane of fracture.
		Conversely, the endocuticle exhibited severe delamination, off-axis ply-splitting, and fiber-pulling away from the fracture surface, indicating the relatively high toughness of the unidirectional \textalpha-chitin organization within the macrofibers \cite{Kamp2010, Kamp2015}.
		Because the exocuticle of weevils is anchored to the endocuticle by cross-linking fibers (see \cite{Kamp2015, Longhai2017}), exocuticular shear-cusp formation in uniaxial tension suggests extension-shear coupling within individual endocuticle laminae, implying that ply-splitting occurred via mode II fracture between macrofibers at high strain \cite{Jones2014, Reddy2004}.
		We hypothesize that intra-laminar extension-shear coupling also yielded off-axis, in-plane resultant forces as a function of lamina orientation angle; mode III shearing then occurred between laminae with opposing in-plane resultant forces, causing the observed inter-ply delamination.
		Tensile failure of the macrofiber laminae would ultimately occur via mixed-mode I/II (transverse tension/intra-laminar shear) fracture due to an increase in applied stress caused by ply-splitting in adjacent laminae \cite{Greenhalgh2009}.

		At the meso-scale, most specimens fractured along a single plane across and between the occipital sulci, which are cuticle invaginations that traverse the entire length of the rostrum \cite{Davis2014, Dennell1942}.
		These sulci increase the volume fraction of exocuticle in the ventral part of the snout and contain large interfaces ideal for void nucleation.
		The exocuticle of the occipital sulci usually displayed shear cusps oriented outward from the center of the invagination, continuing dorsolaterally and ventromedially.
		Ventrally the cusps converged toward a prominent scarp where the crack fronts joined, although in specimens with large cross-sectional areas of endocuticle, this was often obscured by delaminated endocuticular macrofibers.

>>>>

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{gensymb}
\usepackage{textgreek}
\usepackage{xcolor}
\usepackage[colorlinks]{hyperref}
	\hypersetup{colorlinks,
	linkcolor={red!50!black},
	citecolor={blue!60!black},
	urlcolor={blue!40!black}
	}
\graphicspath{{figures/}}
%-------------------------------------------------------------------------------

\begin{document}
	\begin{abstract}
		The acorn weevil snout exhibits remarkable flexibility and toughness that are derived from from the microarchitecture of its exoskeleton.
		Here we characterize modifications to the composite profile of the rostral cuticle that simultaneously enhance the flexibility and fracture toughness of the distal portion of the snout.
		Using Classical Laminate Plate Theory (CLPT) we estimate the effect of these modifications on the elastic behavior of the exoskeleton.
		Additionally, changes in the relative layer thicknesses and orientation angles of layers in the exoskeleton are related to the tensile behavior of the snout in six species of diverse morphology, and we demonstrate that a highly curved rostrum can be completely straightened without structural damage.
		We demonstrate that increased endocuticle thickness is strongly correlated with increased tensile strength in the snout.
		Consequently, snout stiffness is shown to be inversely correlated with fracture toughness.
		Finally, we identify exocuticle rich invaginations of the occipital sutures both as a likely site of crack initiation in tensile failure and as a source of morphological constraint on the evolution of the snout.

@@

	%title and author block
	{\title{Exoskeletal strength and cuticle composite profile of the acorn weevil rostrum}
	\date{\today}

@@

	The exoskeleton of Coleoptera (beetles) is a hierarchically-structured fibrous composite characterized by variously arranged \textalpha-chitin (N-acetylglucosamine) nanofibrils embedded in a heterogeneous protein matrix \cite{Hepburn1973, Jansen2016, Vincent2004}.
	Although \textalpha-chitin is brittle and strongly anisotropic, beetle cuticle is simultaneously rigid and tough due to its unique laminate microstructure (see kamp, vincent for review), which we characterize in detail below \cite{Kamp2010, Kamp2015, Neville1976}.
	Impact-prone areas and exaggerated structures in arthropods generally exhibit cuticle organization that resists deformation and fracture (e.g. \cite{Amini2015, Mccullough2014mech, Mccullough2014struc, Dirks2012, Dirks2013}); however, acorn weevils in the genus \textit{Curculio} Linnaeus, 1758\footnote{
		Pursuant to the International Code of Zoological Nomenclature, the first mention of any specific epithet will include the full genus and species names as a binomen (two part name) followed by the author and date of publication of the name.
		This is not an in-line reference; it is a part of the name itself and refers to a particular species-concept as indicated in the description of the species by that author \cite{ICZN1999}.}
	instead exhibit unusual distal flexibility in an elongate extension of the head called the rostrum (snout) \cite{Toju2005, Jansen2016, Singh2016, Gibson1969}.
	The rostrum is a hollow, strongly curved (over 90$\degree$ in some species), cylindrical, exoskeletal extension of the otherwise nearly-spherical head, which bears at its apex the terminal chewing mouthparts \cite{Morimoto2003, Ting1933, Ting1936, Dennell1942, Gibson1969}.
	This structure is used by the female to feed and to excavate sites for egg-laying (oviposition, see Fig.~\ref{fig::curculio}), and can be repeatedly bent without evident damage despite being composed of the same material as other rigid body parts \cite{Jansen2016, Singh2016, Gibson1969, Toju2005}.
	By maintaining constant pressure on the snout and rotating around the bore-hole, females are able to flex the rostrum into a straightened configuration and produce a linear channel into the fruit; a single female may prepare hundreds of such sites as an adult \cite{Toju2005, Moffett1989, AguirreUribe1978}.

@@

	\begin{figure*}
		\centering
		\includegraphics[width=180mm]{fig1.pdf}
		\caption{\textbf{Morphology and oviposition behavior of female \textit{Curculio}.} 
			\textbf{a}, Lateral habitus image of female \textit{Curculio~sayi} ((Gyllenhal, 1836)) featuring the elongate, strongly curved rostrum.
			\textbf{b}, Lateral view of the head of a female specimen of \textit{Curculio~longinasus}, with major anatomical features indicated.
			\textbf{c}, Illustration of the oviposition behavior of female \textit{Curculio}, proceeding from left to right: a female makes an incision in the host fruit, flexes the head directly over the bore hole using the front legs, then maintains tension on the snout while rotating to excavate a linear channel into the fruit. During this process, the rostrum is bent until completely straight.
		}
		\label{fig::curculio}
	\end{figure*}

	We additionally describe the fracture mechanics of the snout and consider how modification of the cuticle may prevent crack formation during oviposition.
	Based on our findings, we posit that the composite profile of the rostral apex enables the snout to undergo repeated distal flexion while remaining within the elastic limits of the material, mitigating the risk of structural damage, and without evident alteration of the mechanical properties of the individual components of the cuticle across the structure and between species.
	Thus, the flexibility and tensile strength of the rostrum appear to be derived \emph{exclusively} from modification of the composite architecture of the exoskeleton.
	To our knowledge, this is the first time that a modified composite profile has been reported as a means of enhancing structural elasticity in the insect exoskeleton (but see \cite{Matsumura2017}).

	\section{Microstructure of the \textit{Curculio} rostrum}
		In arthropods (including beetles), the exocuticle is comprised of numerous unidirectional laminae of chitin nanofibrils; each layer is the thickness of a single fiber (2-4\,nm) embedded in a proteinaceous matrix \cite{Nikolov2010, Nikolov2011}.
		These layers are stacked at a more or less constant angle to each other, forming a quasi-isotropic laminate known as the Bouligand structure \cite{Blackwell1980, Bouligand1972, Neville1976}. 
		This layout effectively mitigates the strong anisotropy of \textalpha-chitin to yield a versatile building material for the exoskeleton \cite{Vincent1982, Vincent2004, Nikolov2010, Nikolov2011}.		
		Beetle endocuticle, however, is unique among arthropods and is comprised of large (1-5\,{\textmu}m diameter in \textit{Curculio}) unidirectional bundles of chitin, called macrofibers.
		Chitin macrofibers are orthotropic (axial: $E_1~=~8.5\,\text{GPa}$, transverse: $ E_2~=~E_3~=~0.52\,\text{GPa}$ \cite{Jansen2016}) and arranged in unidirectional plies, seen in Figs.~\ref{fig::cuticle},~\ref{fig::profile} \cite{Kamp2010, Kamp2015}.
		Typically, adjacent macrofiber laminae are paired and pseudo-orthogonal (i.e., angled approx. $90\degree$ to each other \cite{Cheng2009}, see Fig.~\ref{fig::profile}), with a constant stacking angle \emph{between} pairs, although other configurations have been observed \cite{Hepburn1973, Kamp2010, Kamp2015, Leopold1992}.
		This geometric sequence of the macrofiber laminae yields an approximately transversely isotropic composite, similar to the Bouligand structure \cite{Kamp2015, Nikolov2010}.
		Notably, the resulting laminate is less rigid than the exocuticle, but exhibits greater toughness because the pseudo-orthogonal plies effectively inhibit crack formation and propagation between successive layers \cite{Kamp2010, Kamp2015, Hepburn1973}.

		\begin{figure*}
			\centering
			\includegraphics[width=180mm]{fig2.pdf}
			\caption{\textbf{Gross divisions of cuticle in the rostrum.}
				\textbf{a}, Scanning electron micrograph of the head capsule, in frontal view, of female \textit{Curculio~sulcatulus} (Casey, 1897), with the rostrum removed.
				\textbf{b}, Magnified view of the junction between the rostrum and head capsule showing the division of the cuticle into two general regions: the exocuticle and the endocuticle.
			}
			\label{fig::cuticle}
		\end{figure*}

		Serial thin sectioning and scanning electron microscopy of fractured \textit{Curculio} specimens revealed that endocuticle in the head capsule fits this general profile, with an angle of approximately $30\degree$ between successive pairs of pseudo-orthogonal plies.
		Additionally, in the head capsule, the thickness of the exocuticle and endocuticle in cross section are nearly equal (typically between 20-30\,{\textmu}m).
		However, we found that the cuticle composite lay-up of the rostral apex is strongly differentiated from the head capsule (Fig.~\ref{fig::profile}).
		Distally the exocuticle is reduced to a thin shell (ca. 5\,{\textmu}m), with the endocuticle thickened to offset this reduction and maintain a constant cuticle thickness (ca. 50\,{\textmu}m total) throughout its length.
		Moreover, the endocuticular macrofibers exhibit no rotation between successive pseudo-orthogonal plies, and are oriented at approximately $\pm45\degree$ to the longitudinal axis of the snout (i.e., an antisymmetric $[\pm45\degree]$ angle-ply laminate).
		We previously identified these modifications to the composite structure of the cuticle within a single species, \textit{C.~longinasus} Chittenden, 1927 \cite{Jansen2016, Singh2016}; however this composite profile has now been uncovered in the rostrum of six additional, phylogenetically disparate, species (detailed in methods), indicating that this is likely a genus-wide trait.
		In all examined species, the portion of the snout between the head capsule and apex of the scrobe exhibits a gradual transition in composite profile along an anterior-posterior gradient.

		\begin{figure*}
			\centering
			\includegraphics[width=180mm]{fig3.pdf}
			\caption{\textbf{Composite profiles of rostral cuticle.}
				\textbf{a}, Scanning electron micrograph of a fractured specimen of \textit{Curculio~humeralis} (Casey, 1897), showing that the cuticle of the rostral apex is organized as a $\pm45\degree$ angle-ply laminate..
				\textbf{b}, Scanning electron micrograph of a fractured specimen of \textit{Curculio~caryae}  (Horn, 1873), showing that the cuticle of the rostral base and head capsule has a roughly $30\degree$ stacking angle between each pair of pseudo-orthogonal plies.
				Semi-thin sections of cuticle from a specimen of \textit{C.~humeralis}, stained with toluidine-blue-borax, demonstrating:
				(\textbf{c}) that the exocuticle of the rostral apex is reduced to a thin shell close to 5\,{\textmu}m thick, with the endocuticle thickened to maintain a constant laminate thickness; and,
				(\textbf{d}) that the exocuticle of the head capsule and base of the snout occupies nearly half of the through-thickness of the cuticle.
			}
			\label{fig::profile}
		\end{figure*}

@@

	\section{Force-controlled loading to fracture}

		\begin{figure*}
			\centering
			\includegraphics[width=180mm]{fig4.pdf}
			\caption{\textbf{Tensile properties of the \textit{Curculio} rostrum.}
				Each plot shows the relationship between two variables as predicted by a phylogenetic linear mixed-effect model, with species as a random effect and a variance-covariance matrix generated from Brownian motion over the preferred phylogeny of Bonal et al. \cite{Bonal2016}.
				The gray regions represent the prediction interval and bootstrapped 95\% confidence interval of the model.
				The estimated fixed effect $\hat{\beta}$ is given, along with the p-value of a t-test assessing whether $\hat{\beta}$ is significantly different from zero.
				Also reported is a generalized marginal $\text{R}^2_{\beta^*}$ for assessing fixed effects (see online Methods for more information).
				In general, we find that increased endocuticle thickness is associated with greater tensile strength, and that stiffness is inversely correlated with toughness.
			}
			\label{fig::tensile}
		\end{figure*}

		To better characterize the failure behavior of the rostrum, we performed tensile testing on the snouts of six \textit{Curculio} species that representing a mixture of closely and distantly related taxa (see Methods) \cite{Hughes2004phylo, Hughes2004eco, Bonal2016, Bonal2011}.
		Each specimen was immersed in $\text{di-H}_2\text{O}$ for 24 hours, to simulate the living tissue (see \cite{Klocke2011}), then subjected to force-controlled, uniaxial loading to fracture at a constant stress rate of $1.0\,\text{gf}\cdot\text{s}^{-1}$ (detailed in Methods).
		In general, the specimens exhibited a non-linear viscoelastic response curve characterized by a sharp increase in stress at higher strains, terminating in brittle fracture \cite{Mihai2017}.
		We postulate that strain hardening occurs as the longitudinal axis of the macrofibers becomes more closely aligned to the cylindrical axis of the rostrum, thereby resisting tension more directly with increasing strain \cite{Munster2013}.

		We examined the correspondence between composite structure and mechanical behavior of the snout using phylogenetic linear mixed effects models (PGLM, Fig.~\ref{fig::tensile}) to account for phylogenetic non-independence in residual variance, with species membership included as a random effect \cite{Revell2010, Felsenstein1985}.
		Our models show that the maximum force sustained at the site of failure was strongly correlated with the cross-sectional area of the endocuticle ($\hat{\beta}=1.28, p<0.0001$), and not the exocuticle ($\hat{\beta}=0.17, p=0.38$), at that site.
		Accordingly, there was a negative correlation between ultimate tensile strength of the specimen and the ratio of exocuticle to endocuticle cross-sectional area at the site of fracture ($\hat{\beta}=-0.36, p=0.017$).
		Although CLPT predicts a positive association between the proportion of exocuticle and stiffness of a generalized cuticle, we found no evidence of correspondence between the cross-sectional properties of the fracture site and the gross behavior of the entire rostrum.
		Because the cross-sectional areas of the cuticle regions vary across the length of the head along an anterior-posterior gradient, it is not possible to correlate measurements from the fracture surface to the properties of the entire snout.

		Instead, we found that the uniaxial elastic modulus (low strain: $E_{low}$) and secant modulus at failure ($E_{sec}$) were inversely correlated with ultimate strain and fracture toughness (see Fig.~\ref{fig::tensile}), implying that stiffer specimens, and by extension, stiffer cuticle profiles, are generally more brittle.
		We also observed a moderate, but not statistically significant, stiffening size-effect with respect to rostral length, contrary to our expectation that a longer, more strongly curved (see \cite{Hughes2004eco, Bonal2011}) snout would require increased flexibility to avoid fracture during oviposition.
		One possible explanation for this trend might be that longer snouts have a longer transition gradient from basal to apical profile to reinforce the junction between the rostrum and head capsule against buckling.
		Thus, Young's modulus of the rostrum would be comparatively higher in these species because of the higher volume fraction of exocuticle.
		From these tests, we infer that the gross elastic behavior of the cuticle is consistent across the genus, and that a single mechanism (i.e. the modified composite profile) confers increased flexibility and tensile strength to the rostral apex in the genus \textit{Curculio}.
		In addition, the endocuticle demonstrably contributes more to rostral tensile strength than the exocuticle, likely because of its organization into large bundles of aligned, anisotropic fibers, leading to a trade-off between rigidity and toughness.
		Consequently, the altered composite profile of the cuticle in the rostral apex makes the rostrum simultaneously more flexible and fracture resistant.

	\section{Load cycling of \textit{Curculio caryae}}

		\begin{figure*}
			\centering
			\includegraphics[width=180mm]{fig5.pdf}
			\caption{\textbf{Fatigue testing of a female C. caryae.}
				\textbf{a}, Shows the fatigue testing setup, wherein the head capsule was fixed to a pedestal and the rostrum was attached to a strip of rip-stop nylon fabric using cyanoacrylate adhesive.
				The specimen was loaded in tension, and not compression, isolating the effect of tension on the fatigue behavior of the rostrum.
				\textbf{b}, An overlay of the pre- and post-fatigue states of the head show a clear effect from repeated proscribed strain; however,
				\textbf{c}, Demonstrates that this effect is apparently not permanent.
			The first two panels to the left correspond to the pre- and post-fatigue states of the rostrum, respecively, but the third panel shows that the head has returned to its original shape after 24 hours soaking in a water/ethanol mixture.
				\textbf{d}, The two plots presented here are the raw force and displacment data from the fatigue test.
				The plot on the right shows clear viscoelastic behavior, indicated by histeresis in the stress-strain response of the specimen, and exhibits a logarithmic decrease stress amplitude over time.
			}
			\label{fig::fatigue}
		\end{figure*}

		To confirm that repeated, proscribed straightening of the rostrum does not result in damage to the cuticle, we performed displacement-controlled fatigue testing on a typical female specimen of \textit{Curculio caryae}, a species that exhibits extreme ($80-90\degree$, Fig.~\ref{fig::fatigue}) rostral curvature \cite{AguirreUribe1978, Gibson1969}.
		The specimen was aligned so that uniaxial tension would induce elongation of the distal portion of the rostrum with minimal off-axis deflection of the un-curved section.
		The strain per cycle was fixed at an amplitude sufficient to completely elongate the snout and generate a tensile load of 1.0\,N  in the straightened configuration (ca. 20\% ultimate strength) at a frequency of $0.33\,\text{Hz}$.
		The test was terminated after a period of two weeks (ca.~400K cycles) when the stress amplitude appeared to reach an asymptotic minimum.		
		The rostrum behaved viscoelastically, indicated by histeresis in the stress-strain relationship during each cycle.
		Strain amplitude decreased logarithmically with cycle number, and the specimen initially appeared to have deformed plastically during the test.
		We believed that this indicated damage to the specimen; however, after cleaning the specimen in a 24 hour wash with ethanol and water, we observed that rostrum returned to its original shape (Fig.~\ref{fig::fatigue}).

@@

	\section{Fractography of test specimens}

		\begin{figure*}
			\centering
			\includegraphics[width=180mm]{fig6.pdf}
			\caption{\textbf{Fractography of the \textit{Curculio} rostrum.}
				\textbf{a}, Light micrograph showing the fracture surface of a tensile tested rostrum, in the species \textit{C.~caryae}, exhibiting a typical failure mode, illustrated in (\textbf{b}).
				The small, black arrows indicate the winding direction of the macrofiber lamina; green arrows and dotted lines indicate the direction of crack.
				Scanning electron micrographs highlight: (\textbf{c}) shear cusp formation, (\textbf{d}) tensile failure and off-axis macrofiber fiber-pulling, (\textbf{e}) interlaminar delamination, (\textbf{f},\textbf{h}), crack formation near invaginated exocuticle (\textbf{g}) ventral crack-front coalescence and shear cusp formation.
			}
			\label{fig::fracture}
		\end{figure*}

		Although it was not always possible to identify void nucleation and crack initiation sites given the complex failure modes evident in the fractured specimens, we observed several patterns characteristic of both the micro-scale behavior of the cuticle and the meso-scale behavior of the rostrum during uniaxial tensile failure (Fig.~\ref{fig::fracture}), which we describe here.

		In transverse view, the exocuticle consistently presented a nearly continuous fracture surface, characteristic of comparatively brittle failure, presumably due to the relatively homogeneous arrangement of \textalpha-chitin laminae in the Bouligand structure \cite{Nikolov2010, Nikolov2011}.
		The exocuticle typically appeared to fracture at lower strains than the endocuticle, with shear-cusp formation evident both at the fracture surface and across exocuticle adjacent to the plane of fracture \cite{Greenhalgh2009}.
		Conversely, the endocuticle exhibited severe delamination, off-axis ply-splitting, and fiber-pulling away from the fracture surface, indicating the relatively high toughness of the unidirectional \textalpha-chitin organization within the macrofibers \cite{Kamp2010, Kamp2015}.
		Because the exocuticle of weevils is anchored to the endocuticle by cross-linking fibers (see \cite{Kamp2015, Longhai2017}), exocuticular shear-cusp formation in uniaxial tension suggests extension-shear coupling within individual endocuticle laminae, implying that ply-splitting occurred via mode II fracture between macrofibers at high strain \cite{Jones2014, Reddy2004}.
		We hypothesize that intra-laminar extension-shear coupling also yielded off-axis, in-plane resultant forces as a function of lamina orientation angle; mode III shearing then occurred between laminae with opposing in-plane resultant forces, causing the observed inter-ply delamination.
		Tensile failure of the macrofiber laminae would ultimately occur via mixed-mode I/II (transverse tension/intra-laminar shear) fracture due to an increase in applied stress caused by ply-splitting in adjacent laminae \cite{Greenhalgh2009}.

		At the meso-scale, most specimens fractured along a single plane across and between the occipital sulci, which are cuticle invaginations that traverse the entire length of the rostrum \cite{Davis2014, Dennell1942}.
		These sulci increase the volume fraction of exocuticle in the ventral part of the snout and contain large interfaces ideal for void nucleation (see Fig.~\ref{fig::fracture}).
		The exocuticle of the occipital sulci usually displayed shear cusps oriented outward from the center of the invagination, continuing dorsolaterally and ventromedially.
		Ventrally the cusps converged toward a prominent scarp where the crack fronts joined, although in specimens with large cross-sectional areas of endocuticle, this was often obscured by delaminated endocuticular macrofibers.
